---
title: "Interactive Maps with Leaflet"
subtitle: "Life Tutorial"
author: "Camilo Pedraza Jimenez, Ray Hossain, Manjiri Satam"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---

<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

First, we install the packages that we need to use.

```{r, include = T}
library(pacman)
pacman::p_load(leaflet, tidyverse, htmltools, terra, sf, scales, glue, readxl, stringi)
```

<h2><strong>General example </strong></h2>
```{r, include = T}
content <- paste(sep = "<br/>",
  "<b><a href='http://www.samurainoodle.com'>Samurai Noodle</a></b>",
  "606 5th Ave. S",
  "Seattle, WA 98138"
)

leaflet() %>% addTiles() %>%
  addPopups(-122.327298, 47.597131, content,
    options = popupOptions(closeButton = FALSE)
  )  
```
<h4><strong>Customized map</strong></h4>
```{r, include = T}
## Create and Customize the Base Map and the Initial View
leaflet() %>%
  setView(lng = 13.3892, lat = 52.5128, zoom = 12) %>% 
  addTiles()
```
<h4><strong>Adding layers</strong></h4>
```{r, include = T}

## Add data layers, Customize
leaflet() %>%
  setView(lng = 13.3892, lat = 52.5128, zoom = 12) %>% 
  addTiles() %>% 
  addCircles(
  lng = 13.3892,   # Longitude of the circle's center
  lat = 52.5128,   # Latitude of the circle's center
  radius = 300,   # Radius of the circle in meters
  color = "blue",   # Color of the circle
)
```
<h4><strong>Adding popups</strong></h4>
```{r, include = T}
# But, we can create popups as well
content <- paste(sep = "<br/>",
                 "<b><a href='https://www.hertie-school.org/en/'>Hertie School</a></b>",
                 "Friedrichstraße 180",
                 "10117 Berlin"
)
  
leaflet() %>% addTiles() %>%
  addPopups(13.389223597194448, 52.51298665365591, content,
            options = popupOptions(closeButton = FALSE)
  )  
```

<h2><strong>Practical exercise</strong></h2>
<h4><strong>Cleaning Data sets</strong></h4>

Now, we import the data sets that we are going to use. The deforestation data was collected by the [Global Forest Watch](https://www.globalforestwatch.org/) and the specific data for Colombia can be found [here](https://gfw.global/478KQ1A). The geographic files for the municipalities of Colombia are shared by [Laboratorio Urbano - Bogotá](https://bogota-laburbano.opendatasoft.com/pages/home/) and they can be accessed [here](https://bogota-laburbano.opendatasoft.com/explore/data set/shapes/export/?location=6,4.85836,-73.2019&basemap=jawg.streets). Although, the files come in different formats, we decided to used the `Shapefile` format for this exercise.

```{r}
deforestation <- read_xlsx("COL.xlsx", 
                           sheet = "Subnational 2 tree cover loss") %>%
  distinct(subnational2, .keep_all = TRUE) %>%
  mutate(subnational2 = toupper(subnational2),
         subnational1 = toupper(subnational1),
         pct_gain = `gain_2000-2020_ha`/area_ha) %>%
  rename(nombre_mpi = subnational2,
         nombre_dpt = subnational1) %>%
  arrange(nombre_mpi)

map <- sf::st_read("shapes/shapes.shp", 
                   stringsAsFactors = FALSE) %>%
  arrange(nombre_mpi)
```

Since we are working with data from Colombia, some of the municipalities for the "deforestation" data set include the accent. Thus, we use the `stri_trans_general()` function from the package `stringi`. However, since the Spanish language also uses the Ñ and the Ü, we do not want ot get rid of those. Therefore, we need to create a function:

```{r, include = T}
# Function to remove accents
remove_accents <- function(input_string) {
  paste(sapply(strsplit(input_string, "")[[1]], function(x) ifelse(x %in% c("Ñ", "Ü"), x, stringi::stri_trans_general(x, "Latin-ASCII"))), collapse = "")
}

# Apply the function to the vector
deforestation$nombre_mpi <- sapply(deforestation$nombre_mpi, remove_accents)
deforestation$nombre_dpt <- sapply(deforestation$nombre_dpt, remove_accents)
```

<h4><strong>Merging Data sets</strong></h4>

After cleaning the data, we need to merge the different files. There are 91 municipalities with no data available in the deforestation data set

```{r, include = T}
#Merging data
deforestation_map <- full_join(map, deforestation, by = c("nombre_mpi", "nombre_dpt"))
save(deforestation_map, file = "deforestation_map.RData")

#No matches
not_m <- map %>%
  anti_join(deforestation, by = c("nombre_mpi", "nombre_dpt")) %>%
  arrange(nombre_mpi)
```

<h4><strong>Getting creative with the map</strong></h4>

Now that we have the data needed for our map, we can create a palette of colors and some tags

```{r, include = T}
#If you have not followed the previous steps, you can import the .RData file
load("deforestation_map.RData")

#The palette of colors is based on a scale of greens that changes based on the gained ha of trees between 2002 and 2020 by municipality
palette <- colorNumeric(palette = "Greens", domain = deforestation_map$`gain_2000-2020_ha`)

#We create some tags by municipality using the glue and the htmltools packages
def_popup <- glue("<strong>{deforestation_map$nombre_mpi}</strong><br />
                    <strong>Department: {deforestation_map$nombre_dpt}</strong><br />
                    Total area (ha): {scales::comma(deforestation_map$area_ha)}<br /> 
                    Gained ha of trees (2002-2020): {scales::comma(deforestation_map$`gain_2000-2020_ha`, accuracy = 1)}<br />
                    Tree cover loss 2022 (ha): {scales::comma(deforestation_map$tc_loss_ha_2022, accuracy = 1)}<br />")  %>%   
  lapply(htmltools::HTML)

#Final map
leaflet() %>%
  #addTiles() %>% #For the default map
  addProviderTiles("CartoDB.Positron") %>% #For one of the options included in the addProviderTiles() function
  #addTiles(urlTemplate = "https://server.arcgisonline.com/ArcGIS/rest/services/Ocean/World_Ocean_Base/MapServer/tile/{z}/{y}/{x}") %>% #For an example of third party tiles
  addPolygons(
    data = deforestation_map,
    fillColor = ~palette(deforestation_map$`gain_2000-2020_ha`),
    label = def_popup,
    fillOpacity = 1,
    color = "grey",
    weight = 1
  )
```

