---
title: "Interactive Maps with Leaflet"
subtitle: "Life Tutorial"
author: "Camilo Pedraza Jimenez, Ray Hossain, Manjiri Satam"
date: "`r format(Sys.time(), '%B %d, %Y | %H:%M:%S | %Z')`"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: tango
    number_sections: no
    theme: cosmo
    toc: no
---
<style>
div.answer {background-color:#f3f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      error = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      comment = NA)
```

First, we install the packages that we need to use.
```{r, include = T}
library(pacman)
pacman::p_load(leaflet, tidyverse, htmltools, terra, sf, scales, glue, readxl, stringi)
```

<h3><strong>Cleaning Datasets</strong></h3>

Now, we import the data sets that we are going to use. The deforestation data was collected by the [Global Forest Watch](https://www.globalforestwatch.org/) and the specific data for Colombia can be found [here](https://gfw.global/478KQ1A). The geographic files for the municipalities of Colombia are shared by [Laboratorio Urbano - Bogotá](https://bogota-laburbano.opendatasoft.com/pages/home/) and they can be accessed [here](https://bogota-laburbano.opendatasoft.com/explore/dataset/shapes/export/?location=6,4.85836,-73.2019&basemap=jawg.streets). Although, the files come in different formats, we decided to used the `Shapefile` format for this exercise. 

```{r}
deforestation <- read_xlsx("C:/Users/cpedr/OneDrive - Hertie School/Semester 3/Intro to DS/Workshop/Exercise/COL.xlsx", 
                           sheet = "Subnational 2 tree cover loss") %>%
  distinct(subnational2, .keep_all = TRUE) %>%
  mutate(subnational2 = toupper(subnational2),
         subnational1 = toupper(subnational1),
         pct_gain = `gain_2000-2020_ha`/area_ha) %>%
  rename(nombre_mpi = subnational2,
         nombre_dpt = subnational1) %>%
  arrange(nombre_mpi)

map <- sf::st_read("C:/Users/cpedr/OneDrive - Hertie School/Semester 3/Intro to DS/Workshop/Exercise/shapes/shapes.shp", 
                   stringsAsFactors = FALSE) %>%
  arrange(nombre_mpi)
```

Since we are working with data from Colombia, some of the municipalities for the "deforestation" dataset include the accent. Thus, we use the `stri_trans_general()` function from the package `stringi`. However, since the Spanish language also uses the Ñ and the Ü, we do not want ot get rid of those. Therefore, we need to create a function:
```{r, include = T}
# Function to remove accents
remove_accents <- function(input_string) {
  paste(sapply(strsplit(input_string, "")[[1]], function(x) ifelse(x %in% c("Ñ", "Ü"), x, stringi::stri_trans_general(x, "Latin-ASCII"))), collapse = "")
}

# Apply the function to the vector
deforestation$nombre_mpi <- sapply(deforestation$nombre_mpi, remove_accents)
deforestation$nombre_dpt <- sapply(deforestation$nombre_dpt, remove_accents)
```

<h3><strong>Merging Datasets</strong></h3>

After cleaning the data, we need to merge the different files. There are 91 municipalities with no data available in the deforestation dataset
```{r, include = T}
#Merging data
deforestation_map <- full_join(map, deforestation, by = c("nombre_mpi", "nombre_dpt"))

#No matches
not_m <- map %>%
  anti_join(deforestation, by = c("nombre_mpi", "nombre_dpt")) %>%
  arrange(nombre_mpi)
```

<h3><strong>Getting creative with the map</strong></h3>

Now that we have the data needed for our map, we can create a palette of colors and some tags

```{r, include = T}
#The palette of colors is based on a scale of greens that changes based on the gained ha of trees between 2002 and 2020 by municipality
palette <- colorNumeric(palette = "Greens", domain = deforestation_map$`gain_2000-2020_ha`)

#We create some tags by municipality using the glue and the htmltools packages
def_popup <- glue("<strong>{deforestation_map$nombre_mpi}</strong><br />
                    <strong>Department: {deforestation_map$nombre_dpt}</strong><br />
                    Total area (ha): {deforestation_map$area_ha}<br /> 
                    Gained ha of trees (2002-2020): {scales::comma(deforestation_map$`gain_2000-2020_ha`, accuracy = 1)}<br />
                    Tree cover loss 2022 (ha): {scales::comma(deforestation_map$tc_loss_ha_2022, accuracy = 1)}<br />")  %>%   
  lapply(htmltools::HTML)

#Final map
leaflet() %>%
  addProviderTiles("OpenStreetMap") %>%
  addPolygons(
    data = deforestation_map,
    fillColor = ~palette(deforestation_map$`gain_2000-2020_ha`),
    label = def_popup,
    fillOpacity = 1,
    color = "grey",
    weight = 1
  )
```



